\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-10-13 doc/flexiblesusy-paper.tex
updated paper to discuss issues with twoloop Higgs corrections a little more and to describe a change I *propose* we make.

\fs allows to add leading two-loop contributions to the CP-even Higgs
tadpoles and self-energies.  For MSSM-like models (with two CP-even
Higgs bosons, one CP-odd Higgs boson, one neutral Goldstone boson)
these corrections can be enabled by setting \code{UseHiggs2LoopMSSM =
  True} in the model file and by defining the effective $\mu$-term
\code{EffectiveMu = \\[Mu]}.  This will add the zero-momentum
corrections of the order $O(y_t^4 + y_b^2 y_t^2 + y_b^4)$, $O(y_t^2
g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$ from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
For NMSSM-like models (with three CP-even Higgs bosons, two CP-odd
Higgs bosons, one neutral Goldstone boson) the two-loop contributions
are enabled by setting \code{UseHiggs2LoopNMSSM = True} and by
defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
  vS / Sqrt[2]}, for example.  This will add the zero-momentum
corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
\cite{Degrassi:2009yq}, plus MSSM-like contributions of the order
$O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$
\cite{Brignole:2001jy,Dedes:2003km}.

@@

\item \code{HighPoleMassPrecision}: This option solves
  Eq.~\eqref{eq:pole-mass-def} exactly by iterating over the momentum
  $p$.  It therefore provides the determination of the pole masses
  with highest precision, but has also the highest execution time.
  Here the one-loop mass matrix $m_{f,1L}^\text{high}$ is diagonalized
  $n$ times, as in the case of \code{MediumPoleMassPrecision},
  resulting in $n$ pole masses $m_{f_k}^\pole$ ($k = 1,\ldots,n$).
  Afterwards, the diagonalization is repeated, this time using the
  calculated pole masses $m_{f_k}^\pole$ for the momentum calculation
  $p^2 = (m_{f_k}^\pole)^2$.  The iteration stops if convergence is
  reached.
\end{itemize}
%
If higher accuracy is required, two-loop corrections to the
self-energies can be added by setting \code{UseHiggs2LoopMSSM = True}
in the MSSM or \code{UseHiggs2LoopNMSSM = True} in the NMSSM in the
model file.  This enables two-loop Higgs FORTRAN routines supplied by
P.~Slavich from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
which add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
$\oabq$, $\oatauq$ and $\oatab$.  Something similar is done for the
NMSSM, but in this case the NMSSM $\oatas$, $\oabas$ pieces come from
\cite{Degrassi:2009yq}, while for $\oatq$, $\oabq$, $\oatauq$ and
$\oatab$ the MSSM pieces are used though it should be understood that
these are not complete in the NMSSM.  Since the Higgs mass is a very
important measurement and the two-loop corrections can be larger than
the current experimental error \cite{Degrassi:2009yq} we recommend to
set these switches to \code{True} in any MSSM-\footnote{In \fs a model
  is MSSM-like if (i) its superpotential is approximately given by the
  MSSM one, (ii) it implements $R$-parity conservation or something
  equivalent, and (iii) contains two CP-even and CP-odd Higgs bosons,
  where one CP-odd Higgs boson may be a Goldstone boson, all with an
  MSSM-like coupling to $t$, $b$ and $\tau$.} and
NMSSM-like\footnote{In \fs a model is NMSSM-like if (i) its
  superpotential is approximately given by the NMSSM one, (ii) it
  implements $R$-parity conservation or something equivalent, and
  (iii) contains three CP-even and CP-odd Higgs bosons, where one or
  two CP-odd Higgs bosons may be Goldstone bosons, all with an
  NMSSM-like coupling to $t$, $b$ and $\tau$\@.  Examples for
  NMSSM-like models are the USSM and the \ESSM.} model.  For models
with a more extended Higgs sector we recommend that the leading log
two-loop corrections are estimated by generalizing those of the MSSM
or NMSSM.

>>>>

\fs allows the user to add leading two-loop contributions to the CP-even Higgs
tadpoles and self-energies.  For MSSM-like models (with two CP-even
Higgs bosons, one CP-odd Higgs boson, one neutral Goldstone boson)
these corrections can be enabled by setting \code{UseHiggs2LoopMSSM =
  True} in the model file and by defining the effective $\mu$-term
\code{EffectiveMu = \\[Mu]}.  This will add the zero-momentum
corrections of the order $O(y_t^4 + y_b^2 y_t^2 + y_b^4)$, $O(y_t^2
g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$ from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
For NMSSM-like models (with three CP-even Higgs bosons, two CP-odd
Higgs bosons, one neutral Goldstone boson) the two-loop contributions
are enabled by setting \code{UseHiggs2LoopNMSSM = True} and by
defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
  vS / Sqrt[2]}, for example.  This will add the zero-momentum
corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
\cite{Degrassi:2009yq}.  If one also wishes to include the MSSM contributions of the order $O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ as well
\cite{Brignole:2001jy,Dedes:2003km} then one can also set both\footnote{Please note that with both of thse options set true it is the NMSSM corrections for $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ which will be applied rather the the MSSM corrections for that order.} \code{UseHiggs2LoopMSSM = True} and \code{UseHiggs2LoopNMSSM = True}.

 In the case of the NMSSM the default option is to have both flags set
 to true as we feel that it is valuable to have these MSSM corrections
 for scenarios where there singlet mixing is very small and in
 particular for cross checks against the MSSM when close to the MSSM
 limit of the model.  However in cases where the singlet mixing is
 large the result at $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ and
 $O(y_\tau^4)$ will not be complete and including these corrections
 could in principle even make the numerical result further away from
 the correct two loop result at that order if there is a cancellation
 with the missing contributions.  So while including such partial
 two-loop corrections is does not change the formal accuracy the user
 should choose whether or not to include these corrections based on
 the physics they study.  Similarly in models that go beyond the NMSSM
 the user must decide based the physics whether on not these
 corrections will give the leading two-loop corrections in their
 model.

@@

\item \code{HighPoleMassPrecision}: This option solves
  Eq.~\eqref{eq:pole-mass-def} exactly by iterating over the momentum
  $p$.  It therefore provides the determination of the pole masses
  with highest precision, but has also the highest execution time.
  Here the one-loop mass matrix $m_{f,1L}^\text{high}$ is diagonalized
  $n$ times, as in the case of \code{MediumPoleMassPrecision},
  resulting in $n$ pole masses $m_{f_k}^\pole$ ($k = 1,\ldots,n$).
  Afterwards, the diagonalization is repeated, this time using the
  calculated pole masses $m_{f_k}^\pole$ for the momentum calculation
  $p^2 = (m_{f_k}^\pole)^2$.  The iteration stops if convergence is
  reached.
\end{itemize}
%
If higher accuracy is required, two-loop corrections to the
self-energies can be added by setting \code{UseHiggs2LoopMSSM = True}
in the MSSM or \code{UseHiggs2LoopNMSSM = True} in the NMSSM in the
model file.  The former enables two-loop Higgs FORTRAN routines
supplied by P.~Slavich from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
which add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
$\oabq$, $\oatauq$ and $\oatab$.  The latter adds corrections
calculated in the NMSSM at $\oatas$, $\oabas$ from
\cite{Degrassi:2009yq}. If wants to include the $\oatq$, $\oabq$,
$\oatauq$ and $\oatab$ the MSSM pieces are used though it NMSSM
corrections at $\oatas$, $\oabas$ along with the MSSM corrections at
$\oatq$, $\oabq$, $\oatauq$ and $\oatab$ then both flags
\code{UseHiggs2LoopMSSM} and \code{UseHiggs2LoopNMSSM} should be set
to \code{True}.  Since the Higgs mass is a very important measurement
and the two-loop corrections can be larger than the current
experimental error \cite{Degrassi:2009yq} we recommend to set these
switches to \code{True} in any MSSM-\footnote{In \fs a model is
  MSSM-like if (i) its superpotential is approximately given by the
  MSSM one, (ii) it implements $R$-parity conservation or something
  equivalent, and (iii) contains two CP-even and CP-odd Higgs bosons,
  where one CP-odd Higgs boson may be a Goldstone boson, all with an
  MSSM-like coupling to $t$, $b$ and $\tau$.} and
NMSSM-like\footnote{In \fs a model is NMSSM-like if (i) its
  superpotential is approximately given by the NMSSM one, (ii) it
  implements $R$-parity conservation or something equivalent, and
  (iii) contains three CP-even and CP-odd Higgs bosons, where one or
  two CP-odd Higgs bosons may be Goldstone bosons, all with an
  NMSSM-like coupling to $t$, $b$ and $\tau$\@.  Examples for
  NMSSM-like models are the USSM and the \ESSM.} model.  However in
such models the user should still consider whether these corrections
are really the leading corrections in the model or if there are other
potentially large two-loop corrections.  For models
with a more extended Higgs sector we recommend that the leading log
two-loop corrections are estimated by generalizing those of the MSSM
or NMSSM.
